comgen-outline

### 2017-09-01 paper/comgen-paper.Rmd
small edits from beginning to case study 1

Generalizations in language are often studied in *generic language* (or, generics): simple statements that convey generalizations about categories [e.g., "Dogs are friendly"; @Carlson1977; @Cohen1999; @Leslie2007; @Nickel2008].
In contrast to *particular statements* which describe individual entities or situations (e.g., "Rufus is friendly") and can be verified through observation, generic statements refer to inherently unobservable categories (e.g., the category of \textsc{dogs}) and convey information that extends beyond the present context, a fact which two-and-a-half-year-olds recognize [@Cimpian2008].
<!-- Additionally, generics are the primary way by which speakers discuss social categories, and thus are key to propagating stereotypes [@Gelman2004; @Rhodes2012; @Leslie2015] and impacting motivation [@Cimpian2010]. -->
Generics are ubiquitous in everyday conversation as well as in child-directed speech [@Gelman2008]. 
Generalizations are much than just generics, however; they can be found in statements about individuals (*habitual language*; e.g., "Mary swims after work.") as well as causality (*causal language*; e.g., "Staring at the sun makes you go blind.").
It is believed that every language can express generalizations [@Behrens2005; @Carlson1995] and that such language is central to the growth of conceptual knowledge [@Gelman2004].

@@

We propose a formal, unified account of generalizations in language that can make precise, quantitative predictions about these simple and ubiquituous utterances.
We do so by first describing the much more well-studied form of generalization: generalization from observations. 
The Bayesian probability calculus describes inductive generalizations in terms of probability, and this is a good model of human generalizations [@Shepard1987; @Tenenbaum2001].
We then formalize how these probabilities are communicated, using a rational model of communication --- the Rational Speech Act theory [@Frank2012; @Goodman2013; @Franke2015; @Goodman2016].
Such models of rational communication describe the inferences listeners can draw from language, and the utterances speakers should say given their state of knowledge and communicative goals. 
Heretofore, the fragments of language formalized in such models have been limited to *particular statements* that convey information about a specific entity or situation [e.g., in hyperbole: "My watch cost a million dollars"; @Kao2014, or vagueness: "John is tall"; @Lassiter2013].
The language of generalizations, however, does not make reference to entities in the world or concrete situations, but instead refers to concepts in the head.

@@

Generalizations are used to make predictions about properties instances that an agent has yet to experience [@HumeTHN]. 
From a very young age, children draw strong inferences about nonobvious object properties from just a few examples [@Baldwin1993; @Gopnik2000;  @XuTenenbaum2007].
As adults we readily predict that a new dog will have fur, drinking a new cup of coffee will cause jitters, and a new day will find me typing.
In each case, we assign a specific exemplar $x$ to a category $k$, and make a prediction that it will have property $f$. 
The entities can be objects (a dog), events (a coffee drink), or more fine grained types (me on a given day).
The properties can vary greatly (e.g., having fur, causing jitters, typing). 
Yet the prediction for a new entity can be described as a *predictive probability*, which we will denote throughout the paper as $h$ and refer to as the **prevalence**: 
$h = P(x \in f \mid x \in k)$, the probability that $x$ will have $f$ given that it is in $k$.^[
We denote prevalence as $h$ to emphasize that it is a latent *hypothesis*, and not necessarily an observable frequency.
]

@@

Studying generalizations from observational evidence can only take us so far, however, in understanding how the mind understands the world or culture accumulates across generations.
Relying upon observations would be particularly problematic to learn, for example, about properties that are difficult to costly to observe (e.g., *starting at the sun makes you go blind*) or events that are statistically rare (e.g., *lightning strikes tall objects*).
For abstract, generalizable knowledge to remain in a culture, it must be able to be faithfully transmitted between individuals and across generations [@Tomasello1999], and language provides simple ways to communicate generalizations.

@@

$\delta_{\denote{u}(h, \theta)}$ is the Kronecker delta function that returns probabilities proportional to $1$ for prevalence values $h$ greater than threshold $\theta$, and $0$ otherwise. 
Logical quantifiers can be described using threshold semantics (i.e., *some* is $h > 0$, all is $h = 1$).
What threshold $\theta$ should be used for the generic generalization?

@@

Given the likelihood "meaning" function defined in Eq. \ref{eq:delta} and the prior distribution over thresholds $\theta$, we can formalize a **listener model** by specifying a prior distribution over the prevalence $P(h)$:

@@

Equation \ref{eq:L0} is a model of a Bayesian listener $L$ who updates her beliefs about the prevalence $h$ according the meaning of an utterance $u$.
$\delta_{\denote{u}(h, \theta)}$ is the likelihood term in Bayes' Rule and represents the literal semantic content of an utterance.
$P(h)$ represents the listener's prior world knowledge concerning the relevant prevalences $h$.
The particular form of the **prevalence prior** $P(h)$ plausibly depends upon the kind of property or event in question and could be structured as a result of deeper conceptual knowledge about properties and events [@Griffiths2005; @Tenenbaum2006; @Kemp2008].
This component of the model provides a potential bridge to conceptual account of generics [e.g., @Leslie2007; @Prasada2006; @Prasada2013].
Critically in our model, this structure is in listener's knowledge of the world $P(h)$ and not in the semantic content of the utterance $u$.

@@

Next, we consider a few classic cases from the literature on generics to demonstrate the kind of inferences our listener model would draw from different generic statements.
The strength of a listener's inference about $h$ (i.e., how prevalent is the feature?) depends in systematic ways on the prior belief distribution over $h$.
In Figure \ref{fig:simulations} (row 2), we show a schematic prior distribution that is bimodal with peaks near 0\% and some high prevalence values; this sort of prior could be over the prevalence of the feature "barks" because many kinds of creatures do not bark (i.e., have $h \sim 0$; e.g., cats) and *for those that do have members that bark*, many of them bark (i.e., have some high $h$).
The fixed threshold model (right column) struggles to differentiate those that generally do not have the feature from those that usually have the feature, returning a posterior distribution very similar to the prior.^[
We model "those that do not generally have the feature" as a Beta distribution with an expected value (mean) close to zero. 
The fact that this component of the prior is not just a Delta function at 0\% reflects the intuitive possibility that some members of a kind who do not generally have the feature (e.g., a cat who barks) could somehow acquire the feature from accidental or transient causes (e.g., a strange genetic mutation).
]
The uncertain threshold model, however, easily rules out the categories that do not generally have the feature and strongly signals the property is widespread in the category.

How does the model handle the puzzling case of "Robins lay eggs" vs. "Robins are female"?
In each case, we must consider the prevalence prior for the feature:  $P(h_\text{lay eggs})$ and $P(h_\text{are female})$.
For being female, no matter want animal category a person might think of, the feature is prevalent in the exact same proprotion (i.e., \%50 of X are female; Figure \ref{fig:simulations} bottom row).
For laying eggs, it really depends: If a person thinks of a falcon, then 50\% of that category lays eggs; if a person thinks of a bear, then 0\% of them lay eggs.
This dependency is reflected in the corresponding $P(h)$, which is a bimodal distribution with peaks near-0\% and near-50\% (Figure \ref{fig:simulations} row 3).
For both priors, the interpretation upon hearing the generalization is the same (i.e., roughly 50\% of the category has the feature; Figure \ref{fig:simulations} rows and 5, middle red panel).
The difference in endorsement comes from the comparison to the prior: For being female, the listener is very likely to arrive at 50\% under her prior, but for laying egges, this is not the case. 
In other words, "Robins are female" has little very informational content as the listener's the posterior is nearly identical with her prior.
"Robins lay eggs", however, has a lot of informational content.

Finally, we look at how the uncertain threshold produces an interpretation for the statement "Mosquitos carry malaria", which intuitively should correspond to "Some mosquitos carry malaria".
The prior $P(h_\text{carry malaria})$ assumes that most creatures do not carry malaria and among those that do, the feature is not very widespread.
Here, we see that listener's posterior upon hearing the generalization (the uncertain threshold) is similar to what she learns about hearing "Some mosquitos carry malaria" (the fixed threshold; Figure \ref{fig:simluations} row 4). 
Thus, a speaker who believes that *very few* mosquitos carry malaria will still endorse "Mosquitos carry malaria".

>>>>

Generalizations in language are often studied in *generic language* (or, generics): simple statements that convey generalizations about categories [e.g., "Dogs are friendly"; @Carlson1977; @Cohen1999; @Leslie2007; @Nickel2008].
Generics are ubiquitous in everyday conversation as well as in child-directed speech [@Gelman2008]. 
In contrast to *particular statements* (e.g., "Rufus is friendly") which describe individual entities or situations and can be verified through observation, generic statements refer to inherently unobservable categories (e.g., the category of \textsc{dogs}) and convey information that extends beyond the present context, a fact which 2-and-a-half-year-olds appreciate [@Cimpian2008].
<!-- Additionally, generics are the primary way by which speakers discuss social categories, and thus are key to propagating stereotypes [@Gelman2004; @Rhodes2012; @Leslie2015] and impacting motivation [@Cimpian2010]. -->
Generics are just one case of generalizations in language, however;  individuals can be described in generalizations (*habitual language*; e.g., "Mary swims after work.") as well as causality (*causal language*; e.g., "Staring at the sun makes you go blind.").
It is believed that every language can express generalizations [@Behrens2005; @Carlson1995] and that such language is central to the growth of conceptual knowledge [@Gelman2004].

@@

We propose a formal, unified account of generalizations in language that can make precise, quantitative predictions about these simple and ubiquituous utterances.
We do so by first reviewing the much more well-studied form of generalization: generalization from observations. 
The Bayesian probability calculus describes inductive generalizations in terms of probability, and this is a good model of human generalization [@Shepard1987; @Tenenbaum2001].
We then formalize how these probabilities are communicated, using a rational model of communication --- the Rational Speech Act theory [@Frank2012; @Goodman2013; @Franke2015; @Goodman2016].
Such models of rational communication describe the inferences listeners can draw from language, and the utterances speakers should say given their state of knowledge and communicative goals. 
Heretofore, the fragments of language formalized in such models have been limited to *particular statements* that convey information about a specific entity or situation [e.g., in hyperbole: "My watch cost a million dollars"; @Kao2014, or vagueness: "John is tall"; @Lassiter2013].
The language of generalizations, however, does not make reference to entities in the world or concrete situations, but rather refer to unobservable concepts in the head.

@@

As adults we readily predict that a new dog will have fur, drinking a new cup of coffee will cause jitters, and a new day will find me typing.
Generalizations are used to make predictions about properties of instances that an agent has yet to experience [@HumeTHN]. 
From a very young age, children draw strong inferences about nonobvious object properties from just a few examples [@Baldwin1993; @Gopnik2000;  @XuTenenbaum2007].
In each case, we assign a specific exemplar $x$ to a category $k$, and make a prediction that it will have property $f$. 
The entities can be objects (a dog), events (a coffee drink), or more fine grained types (me on a given day).
The properties can vary greatly (e.g., having fur, causing jitters, typing). 
Yet the prediction for a new entity can be described as a *predictive probability*, which we will denote throughout the paper as $h$ and refer to as the **prevalence**: 
$h = P(x \in f \mid x \in k)$, the probability that $x$ will have $f$ given that it is in $k$.^[
We denote prevalence as $h$ to emphasize that it is a latent *hypothesis*, and not necessarily an observable frequency.
]

@@

Studying generalizations from observational evidence can only take us so far, however, in understanding how the mind understands the world or culture accumulates across generations.
Relying upon observations would be particularly problematic, for example, to learn about properties that are costly to observe (e.g., *starting at the sun makes you go blind*) or events that are statistically rare (e.g., *lightning strikes tall objects*).
For abstract, generalizable knowledge to remain in a culture, it must be able to be faithfully transmitted between individuals and across generations [@Tomasello1999], and language provides simple ways to communicate generalizations.

@@

Following standard treatment from formal semantics, we represent the semantic content of an utterance by the Kronecker delta function  $\delta_{\denote{u}(h, \theta)}$, which here returns probabilities proportional to $1$ for prevalence values $h$ greater than threshold $\theta$, and $0$ otherwise. 
Logical quantifiers can be described using such a semantics (i.e., *some* is $h > 0$, all is $h = 1$).
What threshold $\theta$ should be used for the generic generalization?

@@

Given the semantic likelihood function defined in Eq. \ref{eq:delta} and the prior distribution over thresholds $\theta$, we can formalize a **listener model** by specifying a prior distribution over the prevalence $P(h)$:

@@

Equation \ref{eq:L0} is a model of a Bayesian listener $L$ who updates her beliefs about the prevalence $h$ according the meaning of an utterance $u$.
$\delta_{\denote{u}(h, \theta)}$ is the likelihood term in Bayes' Rule and represents the literal semantic content of an utterance.
$P(h)$ represents the listener's prior world knowledge concerning the relevant prevalences $h$.
The particular form of the **prevalence prior** $P(h)$ plausibly depends upon the kind of property or event in question and could be structured as a result of deeper conceptual knowledge about properties and events [@Griffiths2005; @Tenenbaum2006; @Kemp2008].
This component of the model provides a potential bridge to conceptual accounts of generics [e.g., @Leslie2007; @Prasada2006; @Prasada2013].
Contrary to these accounts, in our model, conceptual structure is represented in listeners' knowledge of the world $P(h)$ and not in the semantic content of the utterance $u$.

@@

Next, we consider a few classic cases from the literature on generics to demonstrate the kind of inferences our listener model would draw from different generic statements.
The strength of a listener's inference about $h$ (i.e., how prevalent is the feature?) depends in systematic ways on the prior belief distribution over $h$.
In Figure\ \@ref(fig:simulations) (row 2), we show a schematic prior distribution that is bimodal with peaks near 0\% and some high prevalence values; this sort of prior could be over the prevalence of the feature "barks" because many kinds of creatures do not bark (i.e., have $h \sim 0$; e.g., cats) and *for those that do have members that bark*, many of them bark (i.e., have some high $h$).
The fixed threshold model (right column) struggles to differentiate those that generally do not have the feature from those that usually have the feature, returning a posterior distribution very similar to the prior.^[
We model "those that do not generally have the feature" as a Beta distribution with an expected value (mean) close to zero. 
The fact that this component of the prior is not just a Delta function at 0\% reflects the intuitive possibility that some members of a kind who do not generally have the feature (e.g., a cat who barks) could somehow acquire the feature from accidental or transient causes (e.g., a strange genetic mutation).
]
The uncertain threshold model, however, easily rules out the categories that do not generally have the feature and strongly signals the property is widespread in the category.

How does the model handle the puzzling case of "Robins lay eggs" vs. "Robins are female"?
In each case, we must consider the prevalence prior for the feature:  $P(h_\text{lay eggs})$ and $P(h_\text{are female})$.
For being female, no matter want animal category a person might think of, the feature is prevalent in the exact same proprotion (i.e., 50\% of X are female; Figure\ \@ref(fig:simulations) bottom row).
For laying eggs, it really depends: If a person thinks of a falcon, then 50\% of that category lays eggs; if a person thinks of a bear, then 0\% of them lay eggs.
This dependency is reflected in the corresponding $P(h)$, which is a bimodal distribution with peaks near-0\% and near-50\% (Figure\ \@ref(fig:simulations) row 3).
For both priors, the interpretation upon hearing the generalization is the same (i.e., roughly 50\% of the category has the feature; Figure\ \@ref(fig:simulations) rows and 5, middle red panel).
The difference in endorsement comes from the comparison to the prior: For being female, the listener is very likely to arrive at 50\% under her prior, but for laying egges, this is not the case. 
In other words, "Robins are female" has little very informational content as the listener's posterior is nearly identical with her prior (i.e., it doesn't matter if the speaker says the generalization or stays silent).
"Robins lay eggs", however, has a lot of informational content.

Finally, we look at how the uncertain threshold produces an interpretation for the statement "Mosquitos carry malaria", which intuitively should correspond to "Some mosquitos carry malaria".
The prior $P(h_\text{carry malaria})$ assumes that most creatures do not carry malaria and among those that do, the feature is not very widespread.
Here, we see that listener's posterior upon hearing the generalization (the uncertain threshold) is similar to what she learns about hearing "Some mosquitos carry malaria" (the fixed threshold; Figure\ \@ref(fig:simulations) row 4). 
Thus, a speaker who believes that *very few* mosquitos carry malaria will still endorse "Mosquitos carry malaria".
